Lessons learned: creating tutorials teaching Agent-Based Modelling for Archaeologists

### 2024-02-28 paper.Rmd
update paper

---
title: "Lessons learned: creating tutorials teaching Agent-Based Modelling for Archaeologists"
author:
  - name: Ronald M. Visser \orcidlink{0000-0001-6966-1729}
    affil-id: 1
  - name: Annemarie Jutte \orcidlink{0000-0000-0000-0000}
    affil-id: 1
  - name: Doug Rocks-Macqueen \orcidlink{0000-0000-0000-0000}
    affil-id: 2
  - name: Iza Romanowska \orcidlink{0000-0000-0000-0000}
    affil-id: 3
  - name: Tom Brughmans \orcidlink{0000-0000-0000-0000}
    affil-id: 3
  - name: Laura van der Knaap \orcidlink{0000-0000-0000-0000}
    affil-id: 4
  - name: Karsten Lambers \orcidlink{0000-0000-0000-0000}
    affil-id: 4
  - name: Kenneth Aitchison \orcidlink{0000-0000-0000-0000}
    affil-id: 2
affiliations:
  - id: 1
    name: Saxion University of Applied Sciences,
  - id: 2
    name: Landward Research Teoranta.
  - id: 3
    name: Aarhus University.
  - id: 4
    name: Leiden University.
date: "`r format(Sys.time(), '%d %B %Y')`"
geometry:
  - margin=1in
output:
  pdf_document:
    number_sections: yes
    #template: default.tex
bibliography: references.bib
csl: journal-of-computer-applications-in-archaeology.csl
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(ggplot2)
```

@@

The participants of the workshop came from many different countries all over the world.

```{r}
#pre_surveys <- readRDS("data/pre_workshop_surveys.rds")
#ggplot(pre_surveys, aes(x = `What.is.you.nationality....Nationality...Text`)) +
#  geom_bar() +  coord_flip() + xlab("Nationality")

@@

```{r}
#ggplot(pre_surveys, aes(x = `What.is.your.age.`, fill = `What.is.your.gender....Selected.Choice`)) +
#  geom_bar() + xlab("Age") + facet_grid(~Workshop) +
#  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

>>>>

---
title: "Lessons learned: creating tutorials teaching Agent-Based Modelling for Archaeologists"
author:
  - name: Ronald M. Visser
    orcid: 0000-0001-6966-1729
    affil-id: 1
  - name: Annemarie Jutte
    orcid: 0000-0000-0000-0000
    affil-id: 1
  - name: Doug Rocks-Macqueen 
    orcid: 0000-0000-0000-0000
    affil-id: 2
  - name: Iza Romanowska 
    orcid: 0000-0000-0000-0000
    affil-id: 3
  - name: Tom Brughmans
    orcid: 0000-0000-0000-0000
    affil-id: 3
  - name: Laura van der Knaap 
    orcid: 0000-0000-0000-0000
    affil-id: 4
  - name: Karsten Lambers 
    orcid: 0000-0000-0000-0000
    affil-id: 4
  - name: Kenneth Aitchison 
    orcid: 0000-0000-0000-0000
    affil-id: 2
affiliations:
  - id: 1
    name: Saxion University of Applied Sciences.
  - id: 2
    name: Landward Research Teoranta.
  - id: 3
    name: Aarhus University.
  - id: 4
    name: Leiden University.
date: "`r format(Sys.time(), '%d %B %Y')`"
geometry:
  - margin=1in
output:
  pdf_document:
    number_sections: yes
    keep_tex: true
    fig_caption: yes
    latex_engine: pdflatex
  html_document:
    number_sections: yes
  word_document:
    number_sections: yes
bibliography: references.bib
csl: journal-of-computer-applications-in-archaeology.csl
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding,
  output_format = "all") })
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(ggplot2)
library(dplyr)
library(tidyr)
library(forcats)
```

@@

A large proportion of the participants of the workshops gave us information using the survey before the workshops 172 of (...total number...) participants. The respondents came from many different countries all over the world (see Figure \@ref(fig:nationality). The number of female respondents slightly outnumber the male ones and a small group did not share their gender and two were non-binary \@ref(fig:gender-age). The ages of the respondents ranged from below 20 to over 70, reaching people in different age classes, although the majority was aged between 20 and 40.

```{r nationality, fig.cap="The nationality of the participants that filled in the survey.", echo=FALSE}
pre_surveys <- readRDS("data/pre_workshop_surveys.rds")
ggplot(pre_surveys, aes(x = `What.is.you.nationality....Nationality...Text`)) +
  geom_bar() +  coord_flip() + xlab("Nationality")

@@

```{r gender-age, fig.cap="The gender and age of the respondents for each workshop", echo=FALSE}
ggplot(pre_surveys, aes(x = `What.is.your.age.`, fill = `What.is.your.gender....Selected.Choice`)) +
  geom_bar() + labs(x= "Age", y="Count", fill = "Gender") + facet_grid(~Workshop) +   theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

Most of the respondents had several skills with computer applications in archaeology (see Figure @ref(fig:computer-skills)), and many of them did have some knowledge of ABM, although a large group was completely new to the subject. The majority had never applied ABM before participating in the workshops with 14 having applied ABM. It is interesting to note that many respondents did know what kind of software was available for ABM (see Figure \@ref(fig:abm-knowledge)).

```{r, computer-skills, fig.cap="The computer skills of the respondents."}
pre_surveys %>%
  select(`Which.computer.skills.do.you.have....Selected.Choice`) %>%
  separate_rows(`Which.computer.skills.do.you.have....Selected.Choice`, sep = ',\\s*') %>%
  filter(`Which.computer.skills.do.you.have....Selected.Choice`!="") %>%
  group_by(`Which.computer.skills.do.you.have....Selected.Choice`) %>%
  ggplot(aes(x=`Which.computer.skills.do.you.have....Selected.Choice`)) + geom_bar() +
    xlab("Computer skills") + coord_flip()

```{r, abm-knowledge, fig.cap="Respondents knowledge and experience with ABM.", fig.show="hold", out.width="50%"}
pre_surveys %>% 
  filter(`Do.you.know.what.Agent.Based.Modeling.is.`!="Not shared") %>% 
  ggplot(aes(x = `Do.you.know.what.Agent.Based.Modeling.is.`)) +
    geom_bar(na.rm = TRUE) + xlab("I know what ABM is") + facet_grid(~Workshop)
pre_surveys %>%
  select(`Which.software.do.you.know.for.Agent.Based.Modeling....Selected.Choice`) %>%
  filter(`Which.software.do.you.know.for.Agent.Based.Modeling....Selected.Choice`!="") %>%
  separate_rows(Which.software.do.you.know.for.Agent.Based.Modeling....Selected.Choice, sep = ',\\s*') %>%
  group_by(`Which.software.do.you.know.for.Agent.Based.Modeling....Selected.Choice`) %>%
  ggplot(aes(x=`Which.software.do.you.know.for.Agent.Based.Modeling....Selected.Choice`)) + geom_bar() +
  xlab("ABM Software known") + coord_flip()
